# Basic workflow

In this section, we will show a basic workflow for performing an E~max~ model 
analysis for continuous endpoint.

## Setup and load

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(BayesERtools)
library(posterior)
library(tidybayes)
library(here)
library(gt)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
# Enable colored outputs
source(here("R", "cli_color_text.R"))
```

## Load data

The data for the example is included in the BayesERtools package, which contains a simulated data set `d_sim_emax`. The analysis uses the `exposure` variable to predict the continuous outcome `response_1`. The data are illustrated below:

```{r}
#| label: fig-data-look
#| message: FALSE

d_sim_emax |>
  head() |>
  gt() |>
  fmt_number(decimals = 2)

ggplot(d_sim_emax, aes(exposure, response_1)) +
  geom_point() +
  geom_smooth(
    formula = y ~ x,
    method = "loess",
    se = FALSE,
    col = "darkgrey"
  )
```

## Sigmoidal E~max~ model

```{r}
set.seed(1234)

ermod_sigemax <- dev_ermod_emax(
  data = d_sim_emax,
  var_resp = "response_1",
  var_exposure = "exposure",
  gamma_fix = NULL
)

ermod_sigemax
```

## Observation vs model fit

```{r}
#| label: fig-plot-er
#| fig-width: 7
#| fig-height: 5

d_sim_ermod_sigemax <-
  sim_er(ermod_sigemax, output_type = c("median_qi"))

plot_er_gof(d_sim_ermod_sigemax)
```

## Parameter estimates

```{r}
d_draws_sigemax <- as_draws_df(ermod_sigemax)

d_draws_sigemax_summary <-
  summarize_draws(d_draws_sigemax)

ec50_mean <-
  d_draws_sigemax_summary |>
  filter(variable == "ec50") |>
  pull(mean)

d_draws_sigemax_summary |>
  gt() |>
  fmt_number(decimals = 2)
```

## Prediction at a certain concentrations

```{r}
#| label: fig-plot-er-new-conc
#| fig-width: 7
#| fig-height: 5

d_sim_new_conc <-
  sim_er_new_exp(ermod_sigemax,
    exposure_to_sim_vec  = c(5000, 15000, 25000, 35000),
    output_type = c("median_qi"))

d_sim_new_conc |>
  select(-starts_with(".linpred"), -c(.row, .width, .point, .interval)) |>
  gt() |>
  fmt_number(decimals = 2) |>
  tab_header(
    title = md("Prediction at specific concentrations")
  )

d_sim_ermod_sigemax |>
  ggplot(aes(x = exposure, y = response_1)) +
  # Emax model curve
  geom_vline(
    xintercept = ec50_mean,
    linetype = "dashed",
    color = "deepskyblue"
  ) +
  geom_ribbon(
    mapping = aes(
      y = .epred,
      ymin = .epred.lower,
      ymax = .epred.upper
    ),
    alpha = 0.3,
    fill = "deepskyblue"
  ) +
  geom_line(
    mapping = aes(y = .epred),
    color = "deepskyblue"
  ) +
  # Observed data
  geom_point(
    data = d_sim_emax,
    color = "grey"
  ) +
  # Prediction at the specified doses
  geom_point(
    data = d_sim_new_conc,
    mapping = aes(y = .epred),
    color = "tomato",
    size = 3
  ) +
  geom_errorbar(
    data = d_sim_new_conc,
    mapping = aes(
      y = .epred,
      ymin = .epred.lower,
      ymax = .epred.upper
    ),
    width = 1,
    color = "tomato"
  ) +
  coord_cartesian(ylim = c(7, 21)) +
  scale_fill_brewer(palette = "Greys") +
  labs(
    title = "Sigmoidal E~max~ model predictions at new exposure levels",
    caption = paste(
      "vertical dashed line: estimated EC~50~ value",
      "area: 95% credible interval",
      sep = "<br>"
    )
  ) +
  theme(
    plot.title = ggtext::element_markdown(),
    plot.caption = ggtext::element_markdown()
  )
```

