# Covariate modeling with brms

The previous chapter introduced the brms approach to Bayesian E~max~ modeling, with examples provided for hyperbolic and sigmoidal E~max~ models, and considering both continuous and binary outcomes. This chapter extends this by building models that include covariates, and shows examples of model comparison using leave-one-out cross-validation (LOO-CV).

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(brms)
library(posterior)
library(tidybayes)
library(here)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
# Enable colored outputs
source(here("R", "cli_color_text.R"))
```

```{r}
if (require(cmdstanr)) {
  # prefer cmdstanr and cache binaries
  options(
    brms.backend = "cmdstanr", 
    cmdstanr_write_stan_file_dir = here("_brms-cache")
  )
  dir.create(here("_brms-cache"), FALSE)
} else {
  rstan::rstan_options(auto_write = TRUE)
}
```

## Continuous response with covariates

The simulated data set contains three continuous covariates (`cov_a`, `cov_b`, and `cov_c`) that may be related to `response`. The univariate relationships between each covariate and the response are shown below, along with the relationship between `exposure` and `reponse`: 

```{r}
#| label: show-continuous-emax-data
#| fig-height: 8
#| fig-width: 8
#| message: false

d_example_emax_3cov <- read_csv(here("data", "d_example_emax_3cov.csv"))

d_example_emax_3cov |> 
  pivot_longer(
    cols = c(exposure, cov_a, cov_b, cov_c), 
    names_to = "variable",
    values_to = "value"
  ) |> 
  ggplot(aes(value, response)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method = "loess") + 
  facet_wrap(~ variable, scales = "free_x")
```

In the brms framework, the E~max~ function is treated as a structural model and covariates can be placed on any parameter when the model is specified using `brmsformula()`. As an example, the model specified here sets `cov_a`, `cov_b`, and `cov_c` as covariates on the baseline response:

```{r}
covariate_model <- brmsformula(
  response ~ e0 + emax * exposure / (ec50 + exposure), # structural model
  e0   ~ 1 + cov_a + cov_b + cov_c, # covariate model for baseline
  emax ~ 1,                         # covariate model for max response
  ec50 ~ 1,                         # covariate model for EC50
  nl = TRUE
)
```

The measurement model and parameter prior are specified using `brmsfamily()` and `prior()`, and are the same as for the model without covariates:

```{r}
gaussian_measurement <- brmsfamily(
  family = "gaussian", 
  link = "identity"
)

parameter_prior <- c(
  prior(normal(0, 1.5), nlpar = "e0"),
  prior(normal(0, 1.5), nlpar = "emax"),
  prior(normal(2000, 500), nlpar = "ec50", lb = 0)
)
```

To apply this model to the continuous data, pass all three of these to `brm()`:

```{r}
#| label: run-continuous-emax-model-with-covariates
#| message: false
#| results: hide
continuous_covariate_fit <- brm(
  formula = covariate_model, 
  family = gaussian_measurement, 
  data = d_example_emax_3cov, 
  prior = parameter_prior
) 
```

Printing the `continuous_covariate_fit` object provides summary information about the regression coefficients for the covariates and other parameters:

```{r}
continuous_covariate_fit
```

Extending the data visualization used earlier, the model predictions can be plotted as a function of both `exposure` and `cov_a`:

```{r}
cov_a_map <- 
  tibble(
    cov_a = c(2, 5, 8),
    cov_a_group = c("2 (<3.5)", "5 (3.5~6.5)", "8 (≥6.5)")
  )

sim_exposure_cov_a <- 
  continuous_covariate_fit |> 
  epred_draws(newdata = expand_grid(
    exposure = seq(0, 50000, 1000),
    cov_a = c(2, 5, 8), 
    cov_b = 5,
    cov_c = 5
  )) |> 
  median_qi() |> 
  left_join(cov_a_map, by = join_by(cov_a))

d_for_plot <- 
  d_example_emax_3cov |> 
  mutate(
    cov_a_raw = cov_a,
    cov_a = case_when(
      cov_a < 3.5 ~ 2,
      cov_a >= 3.5 & cov_a < 6.5 ~ 5,
      cov_a >= 6.5 ~ 8
    )
  ) |> 
  left_join(cov_a_map, by = join_by(cov_a))

sim_exposure_cov_a |> 
  ggplot(mapping = aes(exposure, .epred)) + 
  geom_path() + 
  geom_ribbon(
    mapping = aes(ymin = .lower, ymax = .upper), 
    alpha = 0.3
  ) +
  geom_point(
    data = d_for_plot,
    mapping = aes(y = response, color = cov_a_raw)
  ) + 
  facet_wrap(~cov_a_group, labeller = label_both) +
  labs(color = "cov_a") + 
  theme(legend.position = "bottom")
```

## Binary response with covariates

Building a covariate model for binary response data follows the same process as for continuous response data. As before, exploratory visualizations are helpful in illustrating the relationships between covariates and the response:

```{r}
#| label: show-binary-emax-data
#| fig-height: 8
#| fig-width: 8
#| message: false

d_example_emax_bin_3cov <- read_csv(here("data", "d_example_emax_bin_3cov.csv"))

d_example_emax_bin_3cov |> 
  pivot_longer(
    cols = c(exposure, cov_a, cov_b, cov_c), 
    names_to = "variable",
    values_to = "value"
  ) |> 
  mutate(response = factor(response)) |> 
  ggplot(aes(response, value)) + 
  geom_violin(draw_quantiles = .5) + 
  facet_wrap(~ variable, scales = "free_y")
```

As in the previous chapter, the only difference between the binary model and the continuous model is the use of the `bernoulli_measurement` model:

```{r}
bernoulli_measurement <- brmsfamily(
  family = "bernoulli", 
  link = "logit"
)
```

The `parameter_prior` and `covariate_model` are the same as before. All three are passed to `brm()`, as shown below: 

```{r}
#| label: run-binary-emax-model-with-covariates
#| message: false
#| results: hide
binary_covariate_fit <- brm(
  formula = covariate_model, 
  family = bernoulli_measurement, 
  data = d_example_emax_bin_3cov, 
  prior = parameter_prior
) 
```

After the sampling finishes, printing the model fit object shows parameter estimates and details about the behavior of the sampler:

```{r}
binary_covariate_fit
```

The data visualization for this model is shown below:

```{r}
cov_a_map <- 
  tibble(
    cov_a = c(2, 5, 8),
    cov_a_group = c("2 (<3.5)", "5 (3.5~6.5)", "8 (≥6.5)")
  )

sim_exposure_cov_a <- 
  binary_covariate_fit |> 
  epred_draws(newdata = expand_grid(
    exposure = seq(0, 50000, 1000),
    cov_a = c(2, 5, 8), 
    cov_b = 5,
    cov_c = 5
  )) |> 
  median_qi() |> 
  left_join(cov_a_map, by = join_by(cov_a))

d_for_plot <- 
  d_example_emax_bin_3cov |> 
  mutate(
    cov_a_raw = cov_a,
    cov_a = case_when(
      cov_a < 3.5 ~ 2,
      cov_a >= 3.5 & cov_a < 6.5 ~ 5,
      cov_a >= 6.5 ~ 8
    )
  ) |> 
  left_join(cov_a_map, by = join_by(cov_a))

sim_exposure_cov_a |> 
  ggplot(mapping = aes(exposure, .epred)) + 
  geom_path() + 
  geom_ribbon(
    mapping = aes(ymin = .lower, ymax = .upper), 
    alpha = 0.3
  ) +
  geom_jitter(
    data = d_for_plot,
    mapping = aes(y = response, color = cov_a_raw),
    width = 0,
    height = .05
  ) + 
  facet_wrap(~cov_a_group, labeller = label_both) +
  labs(color = "cov_a") + 
  theme(legend.position = "bottom")
```

## Setting covariates on other parameters

The previous two examples illustrate covariates placed on the intercept parameter `e0`. It is possible to define covariate models on any parameter within the E~max~ model. 

```{r}
other_covariates <- brmsformula(
  response ~ e0 + emax * exposure / (ec50 + exposure), # structural model
  e0   ~ 1 + cov_a,   # covariate model for baseline
  emax ~ 1 + cov_b,   # covariate model for max response
  ec50 ~ 1,           # covariate model for EC50
  nl = TRUE
)
```

```{r}
#| label: run-other-covariates-emax-model
#| message: false
#| results: hide
other_covariates_fit <- brm(
  formula = other_covariates, 
  family = gaussian_measurement, 
  data = d_example_emax_3cov, 
  prior = parameter_prior
) 
```

Printing the `other_covariates_fit` object provides summary information:

```{r}
other_covariates_fit
```

## Model comparison

The brms package provides a flexible interface for model comparison using LOO-CV and WAIC, using the [loo](http://mc-stan.org/loo/) package internally. One workflow for calling this interface is illustrated in this section, using the three possible E~max~ models as examples: 

```{r}
# no covariates
base_model <- brmsformula(
  response ~ e0 + emax * exposure / (ec50 + exposure),
  e0   ~ 1,
  emax ~ 1,
  ec50 ~ 1,
  nl = TRUE
)

# one predictor on e0
cov_a_model <- brmsformula(
  response ~ e0 + emax * exposure / (ec50 + exposure),
  e0   ~ 1 + cov_a,
  emax ~ 1,
  ec50 ~ 1,
  nl = TRUE
)

# three predictors on e0
cov_abc_model <- brmsformula(
  response ~ e0 + emax * exposure / (ec50 + exposure),
  e0   ~ 1 + cov_a + cov_b + cov_c,
  emax ~ 1,
  ec50 ~ 1,
  nl = TRUE
)
```

In addition to calling `brm()` to estimate regression coefficients, the `add_criterion()` function is called to run the LOO-CV procedure and store the results internally within the brmsfit object:

```{r}
#| results: hide
#| message: false
base_fit <- base_model |> 
  brm(  
    family = gaussian_measurement, 
    data = d_example_emax_3cov, 
    prior = parameter_prior
  ) |> 
  add_criterion("loo")
  
cov_a_fit <- cov_a_model |> 
  brm(
    family = gaussian_measurement, 
    data = d_example_emax_3cov, 
    prior = parameter_prior
  ) |> 
  add_criterion("loo")

cov_abc_fit <- cov_abc_model |> 
  brm(
    family = gaussian_measurement, 
    data = d_example_emax_3cov, 
    prior = parameter_prior
  ) |> 
  add_criterion("loo")
```

To compare models that have LOO criteria information added, use `loo_compare()`:

```{r}
model_comparison <- loo_compare(
  base_fit,
  cov_a_fit,
  cov_abc_fit
)

model_comparison
```

In this example, `cov_a_fit` model outperforms the other two models. 

By default the printed output shows the most important columns, but the return value from `loo_compare()` contains additional information relevant to the model comparison. To view all columns, call the print method with `simplify = FALSE`:

```{r}
print(model_comparison, simplify = FALSE)
```


