# Simulation from fitted model

This page showcase the model simulation using the E~max~ model with no covariate.

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(BayesERtools)
library(posterior)
library(tidybayes)
library(bayesplot)
library(loo)
library(here)
library(gt)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
# Enable colored outputs
source(here("R", "cli_color_text.R"))
```

## Load data

```{r}
#| message: FALSE
d_sim_emax
```

## Fit models

::: {.panel-tabset}

### Sigmoidal E~max~ model

```{r}
set.seed(1234)

ermod_sigemax <- dev_ermod_emax(
  data = d_sim_emax,
  var_resp = "response_1",
  var_exposure = "exposure",
  gamma_fix = NULL
)

ermod_sigemax
```

### Regular E~max~ (h fixed at 1)

Another model without sigmoidal component; will be used when we do
model comparison.

```{r}
set.seed(1234)

ermod_emax <- dev_ermod_emax(
  data = d_sim_emax,
  var_resp = "response_1",
  var_exposure = "exposure",
  gamma_fix = 1
)

ermod_emax
```

:::

# Extrapolation 

::: {.panel-tabset}

## without residual error

This represents uncertainty in the model parameters.

- `p(f(theta)|xnew, yobs)`

```{r}
#| label: fig-extrapolation-1
#| fig-width: 7
#| fig-height: 5

new_conc_vec <- seq(0, 70000, by = 500)

ersim_sigemax <-
  sim_er_new_exp(ermod_sigemax, new_conc_vec)

ersim_sigemax_med_qi <-
  ersim_sigemax |>
  calc_ersim_med_qi(qi_width = c(0.5, 0.95))

ggplot(
  data = ersim_sigemax_med_qi, 
  mapping = aes(x = exposure, y = .epred)
) +
  geom_ribbon(
    data = ersim_sigemax_med_qi |> filter(.width == 0.95),
    mapping = aes(ymin = .epred.lower, ymax = .epred.upper),
    fill = "yellow3", 
    alpha = 0.5
  ) +
  geom_ribbon(
    data = ersim_sigemax_med_qi |> filter(.width == 0.5),
    mapping = aes(ymin = .epred.lower, ymax = .epred.upper),
    fill = "orange1"
  ) +
  geom_line(
    data = ersim_sigemax_med_qi |> filter(.width == 0.5), 
    col = "darkred"
  ) +
  geom_point(
    data = d_sim_emax, 
    mapping = aes(y = response_1)
  ) +
  coord_cartesian(ylim = c(5, 22)) +
  labs(
    x = "Exposure",
    y = "Response",
    title = "Extrapolation (no residual)",
    subtitle = "Represents uncertainty in the model parameters (Credible interval)",
    caption = "95% CI in yellow, 50% CI in orange"
  )
```

## with residual error

This represents uncertainty in the model parameters plus the residual error.

- `p(ynew|xnew, yobs)`

```{r}
#| label: fig-extrapolation-2
#| fig-width: 7
#| fig-height: 5

ggplot(
  data = ersim_sigemax_med_qi, 
  mapping = aes(x = exposure, y = .prediction)
) +
  geom_ribbon(
    data = ersim_sigemax_med_qi |> filter(.width == 0.95),
    mapping = aes(ymin = .prediction.lower, ymax = .prediction.upper),
    fill = "yellow3", 
    alpha = 0.5
  ) +
  geom_ribbon(
    data = ersim_sigemax_med_qi |> filter(.width == 0.5),
    mapping = aes(ymin = .prediction.lower, ymax = .prediction.upper),
    fill = "orange1"
  ) +
  geom_line(
    data = ersim_sigemax_med_qi |> filter(.width == 0.5), 
    col = "darkred"
  ) +
  geom_point(
    data = d_sim_emax, 
    mapping = aes(y = response_1)
  ) +
  coord_cartesian(ylim = c(5, 22)) +
  labs(
    x = "Exposure",
    y = "Response",
    title = "Extrapolation (incl. residual)",
    subtitle = "Represents uncertainty + residual error (Prediction interval)",
    caption = "95% PI in yellow, 50% PI in orange"
  )
```

## Overlay Emax and Sigmoidal Emax

No discernible difference between the two models.

```{r}
#| label: fig-emax-sigemax-compare
#| fig-width: 7
#| fig-height: 5

ersim_emax <-
  sim_er_new_exp(ermod_emax, new_conc_vec)

ersim_emax_med_qi <-
  ersim_emax |>
  calc_ersim_med_qi(qi_width = c(0.5, 0.95))

ggplot(
  data = ersim_sigemax_med_qi, 
  mapping = aes(x = exposure, y = .epred)
) +
  geom_ribbon(
    data = ersim_sigemax_med_qi |> filter(.width == 0.95),
    mapping = aes(ymin = .epred.lower, ymax = .epred.upper),
    fill = "orange1", 
    alpha = 0.5
  ) +
  geom_line(
    data = ersim_sigemax_med_qi |> filter(.width == 0.95), 
    col = "darkred"
  ) +
  geom_ribbon(
    data = ersim_emax_med_qi |> filter(.width == 0.95),
    mapping = aes(ymin = .epred.lower, ymax = .epred.upper),
    fill = "turquoise3", 
    alpha = 0.5
  ) +
  geom_line(
    data = ersim_emax_med_qi |> filter(.width == 0.95), 
    col = "steelblue3"
  ) +
  geom_point(
    data = d_sim_emax, 
    mapping = aes(y = response_1)
  ) +
  coord_cartesian(ylim = c(5, 22)) +
  labs(
    x = "Exposure",
    y = "Response",
    title = "Extrapolation (no residual)",
    subtitle = "Represents uncertainty in the model parameters (Credible interval)",
    caption = "Sigmoidal Emax: Orange, Emax: Blue, 50% CI"
  )
```

:::


