# Model comparison between linear and E~max~

This page showcase how to compare model structures between linear and E~max~
logistic regression models

## Setup and load

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(BayesERtools)
library(loo)
library(here)
library(gt)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
# Enable colored outputs
source(here("R", "cli_color_text.R"))
```


## Load data

```{r}
data(d_sim_binom_cov)

d_sim_binom_cov_2 <-
  d_sim_binom_cov |>
  mutate(
    AUCss_1000 = AUCss / 1000, BAGE_10 = BAGE / 10,
    BWT_10 = BWT / 10, BHBA1C_5 = BHBA1C / 5,
    Dose = glue::glue("{Dose_mg} mg")
  )

# Grade 2+ hypoglycemia
df_er_ae_hgly2 <- d_sim_binom_cov_2 |> filter(AETYPE == "hgly2")

var_resp <- "AEFLAG"
var_exposure <- "AUCss_1000"
```

## Fit model

::: {.panel-tabset}

### Linear logistic regression

```{r}
#| label: fig-bin-lin
#| fig-width: 7
#| fig-height: 5

set.seed(1234)
ermod_bin_hgly2 <- dev_ermod_bin(
  data = df_er_ae_hgly2,
  var_resp = var_resp,
  var_exposure = var_exposure
)
ermod_bin_hgly2
plot_er_gof(ermod_bin_hgly2, var_group = "Dose")
```

### E~max~ logistic regression

```{r}
#| label: fig-bin-emax
#| fig-width: 7
#| fig-height: 5

set.seed(1234)
ermod_bin_emax_hgly2 <- dev_ermod_bin_emax(
  data = df_er_ae_hgly2,
  var_resp = var_resp,
  var_exposure = var_exposure
)
ermod_bin_emax_hgly2
plot_er_gof(ermod_bin_emax_hgly2, var_group = "Dose")
```

:::

## Model comparison

You can perform model comparison based on expected log pointwise predictive density (ELPD).
ELPD is the Bayesian leave-one-out estimate (see ?`loo-glossary`).

Higher ELPD is better, therefore linear logistic regression model appears better than E~max~ model.
However, `elpd_diff` is small and similar to `se_diff` ([see here](https://stats.stackexchange.com/questions/608881/how-to-interpret-elpd-diff-of-bayesian-loo-estimate-in-bayesian-logistic-regress)),
therefore we can consider the difference to be not meaningful.

```{r}
loo_bin_emax_hgly2 <- loo(ermod_bin_emax_hgly2)
loo_bin_hgly2 <- loo(ermod_bin_hgly2)

loo_compare(list(bin_emax_hgly2 = loo_bin_emax_hgly2, bin_hgly2 = loo_bin_hgly2))
```
