# Customize ER plots

This page demonstrates the options available in `plot_er_gof()` and how to 
customize ER plots beyond the built-in options.

See also the [function reference](https://genentech.github.io/BayesERtools/reference/plot_er_gof.html)
for complete documentation.

## Setup

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(BayesERtools)
library(here)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
source(here("R", "cli_color_text.R"))
```

## Data and model

```{r}
data(d_sim_binom_cov)

d_sim_binom_cov_2 <-
  d_sim_binom_cov |>
  mutate(
    AUCss_1000 = AUCss / 1000,
    Dose = glue::glue("{Dose_mg} mg")
  )

df_er_ae_hgly2 <- d_sim_binom_cov_2 |> filter(AETYPE == "hgly2")
```

```{r}
set.seed(1234)
ermod_bin <- dev_ermod_bin(
  data = df_er_ae_hgly2,
  var_resp = "AEFLAG",
  var_exposure = "AUCss_1000"
)
```

## `plot_er_gof()` options

### Default plot

```{r}
#| label: fig-default
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin)
```

### Grouping by dose (`var_group`)
  
Color data points and add boxplot by dose group:

```{r}
#| label: fig-var-group
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin, var_group = "Dose")
```

### Show exposure coefficient CI (`show_coef_exp`)

Display the credible interval for the exposure coefficient:

```{r}
#| label: fig-coef-exp
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin, var_group = "Dose", show_coef_exp = TRUE)
```

Customize the position and size of the coefficient label:

```{r}
#| label: fig-coef-exp-pos
#| fig-width: 7
#| fig-height: 5

plot_er_gof(
  ermod_bin, 
  var_group = "Dose", 
  show_coef_exp = TRUE,
  coef_pos_x = 5.5, coef_pos_y = 0.2, coef_size = 5
)
```

### Number of bins for observed probability (`n_bins`)

Change the number of bins used to summarize observed event rates:

```{r}
#| label: fig-n-bins
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin, var_group = "Dose", n_bins = 6)
```

### Confidence interval widths

Control CI widths for observed data summary (`qi_width_obs`), simulated curve 
(`qi_width_sim`), and coefficient (`qi_width_coef`):

```{r}
#| label: fig-qi-width
#| fig-width: 7
#| fig-height: 5

plot_er_gof(
  ermod_bin, 
  var_group = "Dose",
  show_coef_exp = TRUE,
  qi_width_obs = 0.9,
  qi_width_sim = 0.9,
  qi_width_coef = 0.9
)
```

### Boxplot options

Control boxplot visibility and appearance:

```{r}
#| label: fig-boxplot-options
#| fig-width: 7
#| fig-height: 5

# Explicitly add boxplot without var_group
plot_er_gof(ermod_bin, add_boxplot = TRUE)
```

Adjust boxplot height and show y-axis title:

```{r}
#| label: fig-boxplot-height
#| fig-width: 7
#| fig-height: 5

plot_er_gof(
  ermod_bin, 
  var_group = "Dose",
  boxplot_height = 0.25,
  show_boxplot_y_title = TRUE
)
```

### Caption (`show_caption`)

The caption is shown by default. Hide it with `show_caption = FALSE`:

```{r}
#| label: fig-no-caption
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin, var_group = "Dose", show_caption = FALSE)
```

### Using log scale for exposure

Use `*` instead of `+` to apply scales to all panels in the combined plot:

```{r}
#| label: fig-log-scale
#| fig-width: 7
#| fig-height: 5

plot_er_gof(ermod_bin, var_group = "Dose", show_coef_exp = TRUE) *
  xgxr::xgx_scale_x_log10(guide = guide_axis(minor.ticks = TRUE))
```

## Customize each element of GOF plot

For deeper customization (titles, themes, colors, shapes), use 
`return_components = TRUE` to get modifiable ggplot objects, then recombine 
with `combine_er_components()`.

### Basic workflow

```{r}
#| label: fig-return-components
#| fig-width: 7
#| fig-height: 5

# Get components
comps <- plot_er_gof(ermod_bin, var_group = "Dose", return_components = TRUE)

# Modify the main plot
comps$main <- comps$main +
  labs(title = "Exposure-Response Analysis", x = "AUC (ng·h/mL / 1000)")

# Recombine
combine_er_components(comps)
```

### Customizing themes, colors, and shapes

```{r}
#| label: fig-custom-theme
#| fig-width: 7
#| fig-height: 5

comps <- plot_er_gof(ermod_bin, var_group = "Dose", return_components = TRUE)

# Extract original data for custom point layer
origdata <- extract_data(ermod_bin) |>
  mutate(Dose = glue::glue("{Dose_mg} mg"))

comps$main <- comps$main +
  # Add points with custom shapes by dose
  geom_jitter(
    data = origdata,
    aes(x = AUCss_1000, y = AEFLAG, color = Dose, shape = Dose),
    width = 0, height = 0.05, alpha = 0.7
  ) +
  scale_shape_manual(values = c(16, 17, 15)) +
  scale_color_brewer(palette = "Set1", guide = guide_legend(reverse = TRUE)) +
  labs(title = "Custom Theme with Shapes", color = "Dose", shape = "Dose") +
  theme_minimal()

comps$boxplot <- comps$boxplot +
  scale_fill_brewer(palette = "Pastel1") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()

combine_er_components(comps)
```

### Adjusting heights and caption

```{r}
#| label: fig-combine-options
#| fig-width: 7
#| fig-height: 5

comps <- plot_er_gof(
  ermod_bin, 
  var_group = "Dose", 
  show_caption = TRUE,
  return_components = TRUE
)

comps$main <- comps$main + labs(title = "Adjusted Layout")

# Custom heights and suppress caption
combine_er_components(comps, heights = c(0.7, 0.3), add_caption = FALSE)
```

### Applying log scale to components

```{r}
#| label: fig-log-scale-components
#| fig-width: 7
#| fig-height: 5

comps <- plot_er_gof(ermod_bin, var_group = "Dose", return_components = TRUE)

log_scale <- xgxr::xgx_scale_x_log10(guide = guide_axis(minor.ticks = TRUE))

comps$main <- comps$main + log_scale + labs(title = "Log Scale")
comps$boxplot <- comps$boxplot + log_scale

combine_er_components(comps)
```

## Summary

For complete documentation, see the 
[`plot_er_gof()` reference](https://genentech.github.io/BayesERtools/reference/plot_er_gof.html).

**`plot_er_gof()` options:**

| Option | Description | Default |
|--------|-------------|---------|
| `var_group` | Group data by column | `NULL` |
| `add_boxplot` | Add exposure boxplot | `TRUE` if `var_group` set |
| `boxplot_height` | Relative height of boxplot | `0.15` |
| `show_boxplot_y_title` | Show boxplot y-axis title | `FALSE` |
| `n_bins` | Bins for observed probability | `4` |
| `qi_width_obs` | CI width for observed data | `0.95` |
| `qi_width_sim` | CI width for model curve | `0.95` |
| `show_coef_exp` | Show coefficient CI | `FALSE` |
| `coef_pos_x/y` | Coefficient label position | Auto |
| `coef_size` | Coefficient label size | `4` |
| `qi_width_coef` | CI width for coefficient | `0.95` |
| `show_caption` | Show plot caption | `TRUE` |
| `return_components` | Return components for customization | `FALSE` |

**Tips:**

- Use `*` instead of `+` to apply scales to all panels
- Use `return_components = TRUE` for titles, themes, colors, and shapes
- Components: `$main` (ER plot), `$boxplot`, `$caption`
- Use `combine_er_components()` to recombine after modification
