# Placebo handling

This page focuses on E-R data sets that include a placebo group, and how they can be analyzed using the BayesERtools package. 

```{r}
#| output: FALSE
#| message: FALSE

library(tidyverse)
library(BayesERtools)
library(here)

theme_set(theme_bw(base_size = 12))
```

```{r}
#| include: FALSE
# Enable colored outputs
source(here("R", "cli_color_text.R"))
```

## Simulated data

For the purposes of this chapter, the `d_sim_placebo` data set from BayesERtools is used. It is similar to the `d_sim_emax` data set used in the Emax chapters, but includes data from a placebo group. The exposure-response relationships in this data set are well-captured by a hyperbolic Emax model:

```{r}
d_sim_placebo
```

```{r}
#| label: plot-continuous-data
d_sim_placebo |> 
  pivot_longer(
    cols = c(exp_1, cnt_a, cnt_b, cnt_c), 
    names_to = "variable",
    values_to = "value"
  ) |> 
  ggplot(aes(value, rsp_1)) + 
  geom_point(aes(color = factor(dose))) + 
  geom_smooth(formula = y ~ x, method = "loess") + 
  facet_wrap(~ variable, scales = "free_x")
```

## Univariate models with placebo data

There are several ways the analyst might wish to analyse and visualize E-R data that include a placebo group. This section considers the case in which:

- The placebo data are included when estimating model parameters
- The placebo data are included in the data visualization

### Continuous response

Linear model:

```{r}
#| label: linear-er-placebo-included
set.seed(1234)
ermod_lin <- dev_ermod_lin(
  data = d_sim_placebo,
  var_resp = "rsp_1",
  var_exposure = "exp_1"
)

plot_er(
  ermod_lin, 
  show_orig_data = TRUE,
  options_orig_data = list(var_group = "dose")
)
```

Emax model:

```{r}
#| label: emax-er-placebo-included
set.seed(1234)
ermod_emax <- dev_ermod_emax(
  data = d_sim_placebo,
  var_resp = "rsp_1",
  var_exposure = "exp_1"
)

plot_er(
  ermod_emax, 
  show_orig_data = TRUE,
  options_orig_data = list(var_group = "dose")
)
```

### Binary response

Logistic regression:

```{r}
#| label: binary-er-placebo-included
set.seed(1234)
ermod_bin <- dev_ermod_bin(
  data = d_sim_placebo,
  var_resp = "rsp_2",
  var_exposure = "exp_1"
)

plot_er(
  ermod_bin, 
  show_orig_data = TRUE,
  options_orig_data = list(
    var_group = "dose",
    bin_breaks = c(0, 0.001, 6018, 10265, 16964, 44568) 
  )
)
```

Binary Emax regression:

```{r}
#| label: binary-emax-er-placebo-included
set.seed(1234)
ermod_bin_emax <- dev_ermod_bin_emax(
  data = d_sim_placebo,
  var_resp = "rsp_2",
  var_exposure = "exp_1"
)

plot_er(
  ermod_bin_emax, 
  show_orig_data = TRUE,
  options_orig_data = list(
    var_group = "dose",
    bin_breaks = c(0, 0.001, 6018, 10265, 16964, 44568) 
  )
)
```

